---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gghoneycomb

<!-- badges: start -->
[![R-CMD-check](https://github.com/dleopold/gghoneycomb/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dleopold/gghoneycomb/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/dleopold/gghoneycomb/graph/badge.svg)](https://app.codecov.io/gh/dleopold/gghoneycomb)
<!-- badges: end -->

gghoneycomb is a ggplot2 extension for creating honeycomb waffle charts, where each category forms a contiguous region of hexagonal cells proportional to its share of the data.

## Features

- Each category occupies a single connected region of hexagons, so relative sizes read at a glance
- Three silhouette shapes: `"rect"` (clean grid), `"rounded"` (softened corners), `"organic"` (irregular, grown from a seed)
- Two layout modes: `"compact"` (deterministic, perimeter-minimizing or blocky) and `"free"` (stochastic, controlled by `temperature`)
- Compact styles: `"perimeter"` minimizes region boundary length; `"blocky"` carves rectangular blocks (requires `silhouette = "rect"`)
- Honey-inspired discrete color palette via `scale_fill_honey()` and `scale_color_honey()`
- Minimal void-based theme via `theme_honeycomb()`
- Fully reproducible layouts with the `seed` parameter
- Output rotation in 90-degree increments via `rotation` (0, 90, 180, 270)

## Installation

You can install the development version of gghoneycomb from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("dleopold/gghoneycomb")
```

## Quick Start

```{r quick-start, eval = FALSE}
library(ggplot2)
library(gghoneycomb)

pollen <- data.frame(
  species = c("Clover", "Wildflower", "Lavender", "Sunflower", "Other"),
  count   = c(45, 25, 15, 10, 5)
)

ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(n_cells = 100, silhouette = "rounded", layout = "free", seed = 1) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

Map `fill` to the category variable and `weight` to counts or proportions. `coord_equal()` keeps the hexagons regular. Everything else is optional.

## Gallery

### Compact rectangular (default)

The default settings produce a clean rectangular grid with perimeter-minimizing region growth. No seed needed because the layout is deterministic.

```{r gallery-compact-rect, eval = FALSE}
library(ggplot2)
library(gghoneycomb)

pollen <- data.frame(
  species = c("Clover", "Wildflower", "Lavender", "Sunflower", "Other"),
  count   = c(45, 25, 15, 10, 5)
)

ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(n_cells = 100) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

### Organic silhouette with free layout

`silhouette = "organic"` grows an irregular boundary from a central seed. Pair it with `layout = "free"` and `temperature` to control how loosely the regions spread.

```{r gallery-organic-free, eval = FALSE}
ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(
    n_cells    = 120,
    silhouette = "organic",
    layout     = "free",
    temperature = 0.35,
    seed       = 42
  ) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

### Blocky compact style

`compact_style = "blocky"` carves rectangular blocks instead of minimizing perimeter. It requires `silhouette = "rect"` and produces a treemap-like appearance.

```{r gallery-blocky, eval = FALSE}
ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(
    n_cells       = 100,
    silhouette    = "rect",
    layout        = "compact",
    compact_style = "blocky"
  ) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

### Rotated layout

`rotation` rotates the rendered honeycomb in 90-degree increments without changing the allocation. Useful for switching between portrait and landscape orientations.

```{r gallery-rotated, eval = FALSE}
ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(
    n_cells    = 100,
    silhouette = "rect",
    rotation   = 90,
    seed       = 1
  ) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

### Proportions input with reversed palette

If your data is already in proportions, set `values_are = "proportions"`. Pass `direction = -1` to `scale_fill_honey()` to reverse the palette order.

```{r gallery-proportions, eval = FALSE}
pollen_prop <- data.frame(
  species = c("Clover", "Wildflower", "Lavender", "Sunflower", "Other"),
  prop    = c(0.45, 0.25, 0.15, 0.10, 0.05)
)

ggplot(pollen_prop, aes(fill = species, weight = prop)) +
  geom_honeycomb(
    values_are = "proportions",
    n_cells    = 100,
    seed       = 7
  ) +
  scale_fill_honey(direction = -1) +
  theme_honeycomb() +
  coord_equal()
```

## Learn More

Full documentation and rendered examples live at <https://dleopold.github.io/gghoneycomb/>.

The Getting Started vignette walks through silhouette shapes, layout modes, cell count selection, and proportions input in detail:

``` r
vignette("gghoneycomb")
```

## Development Workflow (renv)

This package uses [renv](https://rstudio.github.io/renv/) for local dependency management. The `renv.lock` file is **not committed** to version control; CI workflows install dependencies directly from `DESCRIPTION`.

### Setup

```r
# Activate renv and install dependencies
renv::restore()
```

### Adding Dependencies

```r
# Add a runtime dependency
usethis::use_package("newpkg")

# Add a development dependency
usethis::use_package("devpkg", type = "Suggests")

# Update the lockfile (includes dev deps)
renv::snapshot(dev = TRUE)
```

### Keeping in Sync

```r
# Check status
renv::status()

# Update lockfile after DESCRIPTION changes
renv::snapshot(dev = TRUE)

# Restore from lockfile
renv::restore()
```
