---
title: "Getting Started with gghoneycomb"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with gghoneycomb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to create honeycomb plots with gghoneycomb.

## Example: Pollen composition

The honeycomb layout is useful for showing proportional abundance of
categories. In this example, we visualize the pollen composition of a
honey sample.

```{r setup}
library(ggplot2)
library(gghoneycomb)
```

```{r basic-honeycomb}
pollen <- data.frame(
  species = c("Clover", "Wildflower", "Lavender", "Sunflower", "Other"),
  count = c(45, 25, 15, 10, 5)
)

ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(n_cells = 80, compaction = 0.9, seed = 123) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

## Choosing the number of cells

If you do not provide `n_cells`, gghoneycomb chooses a value based on the
number of categories and the smallest proportion. As a rule of thumb, cell
counts between 50 and 500 tend to work well for most plots; higher values
increase detail but can make small regions harder to see.

```{r auto-ncells}
ggplot(pollen, aes(fill = species, weight = count)) +
  geom_honeycomb(compaction = 1) +
  scale_fill_honey() +
  theme_honeycomb() +
  coord_equal()
```

## Using proportions explicitly

If you already have proportions, set `values_are = "proportions"` and pass
the values via the `weight` aesthetic.

```{r proportions}
pollen_prop <- data.frame(
  species = c("Clover", "Wildflower", "Lavender", "Sunflower", "Other"),
  prop = c(0.45, 0.25, 0.15, 0.10, 0.05)
)

ggplot(pollen_prop, aes(fill = species, weight = prop)) +
  geom_honeycomb(values_are = "proportions", n_cells = 80, seed = 123) +
  scale_fill_honey(direction = -1) +
  theme_honeycomb() +
  coord_equal()
```
